\begin{abstract}
 

A rudimentary problem in post-verification debugging of digital circuits
 is identifying a potential single fix rectification point and applying a correction. 
 Traditional debugging techniques involve incremental 
 SAT-solving or simulation based validation, while most of the modern
 debugging approaches use concepts from symbolic computer algebra. 
 Due to their bit-blasting nature, the conventional methods have been 
 shown to be exceedingly resource intensive for complex arithmetic 
 circuits, whereas the existing symbolic computer algebra debugging techniques 
 fail to arrive at a conclusive solution on non-standard circuit implementations. 
 To overcome these limitations, this paper presents an automated debugging 
 technique using concepts from 
 symbolic computer algebra to diagnose a multi-error logic circuit for a possible 
 single fix rectification. Starting from an equivalence checking setup modeled as 
 a polynomial ideal membership test in commutative algebra, we analyze the ideal 
 membership residue to identify potential single fix rectifiable locations. Once the 
 possible rectification nets have been identified, we use nullstellensatz principles 
 to verify if indeed a single fix rectification can be applied at any of these locations.
 If a single fix rectification exists, we model the net as an unknown component 
 and formulate the correction as a quantification procedure. 
 We demonstrate the efficacy of our approach through experiments performed
 on various finite field arithmetic circuits.  
 
\end{abstract}
